section_BEGIN
\begin{zsection}
  \SECTION\ fib\_multenv\_postchecking\_call\_error \parents\ circus\_toolkit
\end{zsection}
channel_decl
\begin{circus}
    \circchannel\ out: \nat
\end{circus}
basic_process_BEGIN
\begin{circus}
    \circprocess\ FibStateMultipleEnv ~~\circdef~~ \circbegin
\end{circus}
process_state_and_init
\begin{circusaction}
    \circstate\ FibState ~~==~~ [~ x, y: \nat ~]  \land [ z: \nat ]\\
    InitFibState ~~==~~ [~ FibState~' | x' = y' = 1 ~]
\end{circusaction}
first_action
\begin{circusaction}
    InitFib ~~\circdef~~ out~!1 \then out~!1 \then InitFibState
\end{circusaction}
schema_action
\begin{schema}{OutFibOp}
    \Delta FibState; next!: \nat
\where
    next! = y' = x + y \land x' = y
\end{schema}
recursive_action

Schema expressions transformed into actions, usually happening as action calls,
are allowed, but the user is warned about their innadequacy. They are innadequate
because we are ``casting'' a Z expression into a \Circus\ action, and scoping rules 
for those are different.  
%
\begin{circusaction}
    OutFib ~~\circdef~~ \circmu\ X \circspot \\
    	\t1 (~ \circvar\ next: \nat \circspot \\
	   \t2 OutFibOp \circseq out~!next \then X ~)
\end{circusaction}
%
A Z (schema) expression is self contained: its declarations 
do not need be into scope, whereas a \Circus\ action requires all variables being 
declared to be within the process state scope. Thus, if the schema expression just
mentions $\Delta FibState$, then the warning is harmless, the casting is safe, and 
no error occurs. 

On the othe hand, if input/output or extra variables (beyond the ones within the 
process state's scope) appear, as is the case in $OutFibOp$ (with $next!$),
$next!$ must be into scope, which is the case at the point of call, but not the case 
when $OutFibOp$ is declared. At post-checking time, all calls are re-typechecked a 
second time to enable mutual recursive calls. And at that point, $OutFibOp$ will be
added for post checking cast as a CallAction, which has a type error because $next!$
is not within scope at the declaration time. 

To solve the problem, one just need to add a type-safe cast around $OutFibOp$,
as $\lschexpract OutFibOp \rschexpract$, so that it does not enter the post-checking 
for its declaration as an action, but as a schema expression. Furthermore, this extra
bracket is useful for having a more regular (and unambigous) \Circus\ grammar. 
At the typechecker, we leave room for the unsafe typecase to minimise backward compatibility 
problems with the old \Circus\ grammar, which did not mentioned these needed special brackets.

main_action
\begin{circusaction}
    \circspot InitFib ~\circseq~ OutFib
\end{circusaction}
before_process_END
\begin{circus}
    \circend
\end{circus}
after_process_END
section_END 
