A block-structured symbol table taken from Chapter 2
of Ian Hayes' specification case studies book.

\begin{zed}
[SYM, VAL]
\end{zed}

\begin{schema}{\Delta ST}
 st, st' : SYM \pfun VAL \\
\end{schema}

\begin{schema}{\Xi ST}
  \Delta ST
\where
  st = st'
\end{schema}

\begin{schema}{LookUp}
  \Xi ST \\
  s?:SYM \\
  v!:VAL
\where
  s? \in \dom(st) \\
  v! = st(s?)
\end{schema}

\begin{schema}{Add}
  \Delta ST \\
  s?:SYM \\
  v?: VAL
\where
  s? \notin \dom(st) \\
  st' = st \cup \{ s? \mapsto v? \}
\end{schema}

\begin{schema}{Replace}
  \Delta ST \\
  s?: SYM \\
  v?: VAL
\where
  s? \in \dom(st) \\
  st' = st \oplus \{ s? \mapsto v? \}
\end{schema}

\begin{schema}{Delete}
  \Delta ST \\
  s?: SYM
\where
  s? \in \dom(st) \\
  st' = \{ s? \} \ndres st
\end{schema}

%\begin{theorem}{AddDelete}
%\vdash? \forall \Delta ST; s?: SYM; v?:VAL @ Add \semi Delete \implies st = st'
%\end{theorem}

\begin{zed}
  Report ::= OK
                  | Symbol\_not\_present
		  | Symbol\_present
		  | Not\_within\_any\_block
		  | Symbol\_not\_found
\end{zed}

\begin{schema}{NotPresent}
  \Xi ST \\
  s?: SYM \\
  rep!: Report
\where
  s? \notin \dom(st) \\
  rep! = Symbol\_not\_present
\end{schema}

\begin{schema}{Present}
  \Xi ST \\
  s?: SYM \\
  rep!: Report
\where
  s? \in \dom(st) \\
  rep! = Symbol\_present
\end{schema}

\begin{zed}
Success == [rep!: Report | rep! = OK] \\
STLookUp == (LookUp \land Success) \lor NotPresent \\
STAdd == (Add \land Success) \lor Present \\
STReplace == (Replace \land Success) \lor NotPresent \\
STDelete == (Delete \land Success) \lor NotPresent
\end{zed}

%%Zinword \distro {\oplus}/
%\begin{axdef}
%  \distro: \seq ST \fun ST
%\where
%  \forall s: ST; ss,tt: \seq ST @
%    \distro \langle\rangle = \emptyset \land
%    \distro \langle s \rangle = s \land
%    \distro(ss \cat tt) = \distro(ss) \oplus \distro(tt)
%\end{axdef}

\begin{axdef}
  \distro == \lambda s:\seq (SYM \pfun VAL) @
      \{sym: SYM; val: VAL | (\exists index: \nat @ sym \mapsto val \in s~index \land \forall i:\nat | i > index @ sym \notin \dom(s~i))@ sym \mapsto val \}
\end{axdef}

\begin{zed}
  BST == \seq (SYM \pfun VAL) \\
  \Delta BST == [bst, bst' : BST] \\
  \Xi BST == [\Delta BST | bst' = bst] \\
\end{zed}

\begin{schema}{BSearch0}
  \Xi BST \\
  s?: SYM \\
  v!: VAL
\where
  s? \in \dom(\distro bst) \land v! = (\distro bst)(s?)
\end{schema}

\begin{schema}{BStart0}
  \Delta BST
\where
  bst' = bst \cat \langle \emptyset \rangle
\end{schema}

\begin{schema}{BEnd0}
  \Delta BST
\where
  bst \neq \langle\rangle \land bst' = front(bst)
\end{schema}

\begin{schema}{BReplace00}
  \Delta BST \\
  s?: SYM \\
  v?: VAL \\
  rep!: Report \\
  \Delta ST
\where
  bst \neq \langle\rangle \\
  front(bst') = front(bst) \\
  st = last(bst) \\
  st' = last(bst') \\
  ((s? \in \dom(st) \land st' = st \oplus \{ s? \mapsto v? \} \land rep! = OK)
    \lor
    (s? \notin \dom(st) \land st' = st \land rep! = Symbol\_not\_present))
\end{schema}

\begin{schema}{\Phi BST}
  \Delta BST \\
  \Delta ST
\where
  bst \neq \langle\rangle = front(bst) \\
  st = last(bst) \\
  st' = last(bst')
\end{schema}

\begin{zed}
  BLookUp0 == STLookUp \land \Phi BST \\
  BAdd0 == STAdd \land \Phi BST \\
  BReplace0 == STReplace \land \Phi BST \\
  BDelete0 == STDelete \land \Phi BST \\

  BLookUp1 == \exists \Delta ST @ BLookUp0 \\
  BAdd1 == \exists \Delta ST @ BAdd0 \\
  BReplace1 == \exists \Delta ST @ BReplace0 \\
  BDelete1 == \exists \Delta ST @ BDelete0
\end{zed}

\begin{schema}{Empty}
  \Xi BST \\
  rep!: Report
\where
  bst = \langle\rangle \land rep! = Not\_within\_any\_block
\end{schema}

\begin{schema}{NotFound}
  \Xi BST \\
  s?: SYM \\
  rep!: Report
\where
  bst \neq \langle\rangle \\
  s? \notin \dom(\distro bst) \\
  rep! = Symbol\_not\_found
\end{schema}

\begin{zed}
  BSearch == (BSearch0 \land Success) \lor NotFound \lor Empty \\
  BStart == BStart0 \land Success \\
  BEnd == (BEnd0 \land Success) \lor Empty \\
  BLookUp == BLookUp1 \lor Empty \\
  BAdd == BAdd1 \lor Empty \\
  BReplace == BReplace1 \lor Empty \\
  BDelete == BDelete1 \lor Empty
\end{zed}
