\begin{zsection}
  \SECTION mtr \parents oz\_toolkit, standard\_toolkit
\end{zsection}

\zedsize{\small}
\begin{class}{Clock}
\also
\begin{state}
date:  \nat
\end{state}
\begin{init}
date' = 0\end{init}
\begin{schema}{getDate}
today!:  \nat  
\where
today! = date
\end{schema}
\begin{schema}{nextDay}
\Delta (date)
\where
date' = date + 1
\end{schema} 
\end{class} 

\begin{class}{FareDataBase}[STATION]
\also
\begin{state}
fare: STATION \cross STATION \pfun  \nat  
\end{state} \\ 
\begin{init}
fare' = \emptyset\end{init} \\ 
\begin{schema}{addFare}
\Delta (fare)\\
start?: STATION \\ 
finish?: STATION \\ 
fare?:  \nat  
\where
(start?, finish?) \notin  \dom fare \\ 
fare' = fare \cup \{(start?, finish?) \mapsto fare?\} \\ 
\end{schema} \\ 
\begin{schema}{updateFare}
\Delta (fare)\\
start?: STATION \\ 
finish?: STATION \\ 
newfare?:  \nat  
\where
 (start?, finish?) \in  \dom fare \\ 
fare' = fare \oplus \{(start?, finish?) \mapsto newfare?\} \\ 
\end{schema} \\ 
\begin{schema}{checkFare}
start?: STATION \\ 
finish?: STATION \\ 
fare!:  \nat  
\where
 (start?, finish?) \in  \dom fare \\ 
fare! = fare(start?, finish?)
\end{schema} 
\end{class} 


\begin{class}{Tickets}[TICKET, STATION]
\also
\begin{zed}
TYPE ::= single | multi | season \\
MultiValue == \{50, 100\} \\ 
SeasonDuration == \{7, 31, 365\}
\end{zed}
\begin{state}
singles:  \power TICKET \\ 
multis:  \power TICKET \\ 
seasons:  \power TICKET \\ 
value: TICKET \pfun  \nat  \\ 
expiry: TICKET \pfun  \nat  \\ 
entrypoint: TICKET \pfun STATION 
\where
 \dom value = (singles \cup multis \cup seasons) \\ 
\forall t: (singles \cup multis \cup seasons) @ \# (\{t\} \dres
entrypoint) \leq 1\end{state} \\ 
\begin{init}
expiry' = \emptyset \\ 
singles' = \emptyset \\ 
multis' = \emptyset \\ 
seasons' = \emptyset \\ 
value' = \emptyset \\ 
entrypoint' = \emptyset
\end{init} \\ 
\begin{schema}{issue}
\Delta (expiry, singles, multi, seasons, value)\\
ticket?: TICKET \\ 
today?:  \nat  \\ 
type?: TYPE \\ 
value?:  \nat  
\where
 ticket? \notin (singles \cup multis \cup seasons) \land\\
       (type? = multi \implies value? \in MultiValue) \land\\
       (type? = season \implies value? \in SeasonDuration)\\
type? = single \implies \\
\t1 expiry' = expiry \cup \{ticket? \mapsto today?\} \land \\ 
\t1 singles' = singles \cup \{ticket?\} \land \\ 
\t1 value' = value \cup \{ticket? \mapsto value?\} \land\\ 
\t1 multi' = multi \land seasons' = seasons\\
type? = multi \implies\\
\t1 expiry' = expiry \cup \{ticket? \mapsto (today? + 730)\} \land \\ 
\t1 value' = value \cup \{ticket? \mapsto value?\} \land \\ 
\t1 multis' = multis \cup \{ticket?\} \land \\ 
\t1 singles' = singles \land seasons' = seasons\\
type? = season \implies\\
\t1 expiry' = expiry \cup \{ticket? \mapsto today? + value?\} \land \\ 
\t1 value' = value \cup \{ticket? \mapsto 1\} \land \\ 
\t1 seasons' = seasons \cup \{ticket?\} \land \\ 
\t1 singles' = singles \land  multi' = multi
\end{schema}
\\ 
\begin{schema}{reIssue}
\Delta (expiry, value, entrypoint)\\
ticket?: TICKET \\ 
today?:  \nat  \\ 
value?:  \nat  
\where
 ticket? \in (singles \cup multis \cup seasons) \land\\
     (type? = multi \implies value? \in MultiValue) \land\\
     (type? = season \implies value? \in SeasonDuration)\\
ticket? \in singles \implies\\
\t1 expiry' = expiry \oplus \{ticket? \mapsto today?\} \land \\ 
\t1 value' = value \oplus \{ticket? \mapsto value?\}\\ 
ticket? \in multis \implies\\
\t1 expiry' = expiry \oplus  \{ticket? \mapsto (today? + 730)\} \land \\ 
\t1 value' = value \oplus \{ticket? \mapsto value?\} \\ 
ticket? \in season \implies\\
\t1 expiry' = expiry \oplus \{ticket? \mapsto today? + value?\} \land \\ 
\t1 value' = value \oplus \{ticket? \mapsto 1\}\\
entrypoint' = \{ticket?\} \ndres entrypoint
\end{schema} \\ 
\begin{schema}{enterStation}
\Delta (entrypoint)\\
ticket?: TICKET \\ 
station?: STATION \\ 
today?:  \nat  
\where
 ticket? \in (singles \cup multis \cup seasons) \land \\ 
\t2 value(ticket?) \neq 0 \land \\ 
\t2 expiry(ticket?) \geq today? \land \\ 
\t2 \{ticket?\} \dres entrypoint = \emptyset \\ 
entrypoint' = entrypoint \cup \{ticket? \mapsto station?\} \\ 
\end{schema} \\ 
\begin{schema}{entry}
ticket?: TICKET \\ 
station!: STATION 
\where
 ticket? \in  \dom entrypoint \\ 
station! = entrypoint(ticket?) \\ 
\end{schema} \\ 
\begin{schema}{exitStation}
\Delta (entrypoint, value)\\
fare?:  \nat  \\ 
ticket?: TICKET \\ 
start!: STATION 
\where
 ticket? \in (singles \cup multis \cup seasons) \land \\ 
ticket? \in  \dom entrypoint \\ 
ticket? \in singles \implies\\
\t1 fare? \leq value(ticket?) \land \\ 
\t1 value' = value \oplus \{ticket? \mapsto 0\}\\
ticket? \in multis \implies\\ 
\t1 value' = value \oplus \{ticket? \mapsto max \{0, value(ticket?) - fare\}\}\\ 
ticket? \in season \implies\\
\t1 value' = value\\
start! = entrypoint(ticket?) \\ 
entrypoint' = \{ticket?\} \ndres entrypoint
\end{schema} 
\end{class} 
