
\documentclass{article}

\usepackage{czt}
\usepackage[color]{circus}

\begin{document}

\title{Special \LaTeX\ symbols within \Circus\ tools}
\author{Leo Freitas}
\date{April 2008}

\maketitle

The parser accepts some extra commands useful for \LaTeX\ typesetting.
They are:
\begin{itemize}
   \item \verb'\circblockbegin' and \verb'\circblockend' --- parsed as hard space
   \item \verb'\circminipagebegin' and \verb'\circminipageend' --- parsed as hard space
   \item \verb'\circnote{whatever text without \ or \\ in it}' --- parsed as comment
   \item \verb'\circtoolsoff' and \verb'\circtoolson' --- parsed as comment
   \item \verb'\znewpage' and \verb'\zbreak' --- parsed as new lines
\end{itemize}
%
Their meaning are self-evident: it switches block indentation and
mini-page configuration;~allows the user to add commentary text
within formal material;~enables the user to switch on/off the
Circus tools within parts of the text;~and breaks page or formulae
accordingly.

The commands \verb'\circblockbegin' and \verb'\circblockend' can appear anywhere
a \verb'\begin{block}/\end{block}'
can appear (\textit{i.e.,} within mathematical formulae). And similarly for mini page.
The command \verb'\circnote{...}' is intended to appear within Z or Circus environments,
but could indeed appear anywhere. The commands \verb'\znewpage' and \verb'\zbreak' can
appear only within a Z or Circus environment. Finally, the commands
\verb'\circtoolsoff' and \verb'\circtoolson' can appear only where ever narrative text
can (\textit{i.e.,} they cannot appear within Z or Circus environments.

These commands are parsed as either hard space or comments.
That is, their UNICODE representation is either of a \verb'~' (hard space)
or just ignored (\textit{i.e.,} don't appear at all) as they are comments.
This has the obvious consequence that translation among markups is no longer
injective. This is already the case in Standard Z: \verb'\t1 ... \t9' is
also translated as hard space and would not translate back to \LaTeX\ as
the appropriate tabbing commands.

\begin{zsection}
   \SECTION special\_symbols \parents circus\_prelude
\end{zsection}

\begin{circus}
   \circchannel\ c \\
   \circnote{This is a chance to add comments to the user}
    \circprocess\ P \circdef \circbegin \\
       C \circdef c \then \Skip \\
       \znewpage
    \circspot \Skip \\
    \zbreak
    \circend
\end{circus}

\circtoolsoff

\begin{circus}
   \circchannel should\_not\_be\_processed : \nat
\end{circus}

\circtoolson

\begin{circus}
   \circchannel should\_be\_processed : \nat
\end{circus}

\end{document}