
\begin{zsection}
   \SECTION  bbook \parents standard\_toolkit
\end{zsection}

\begin{zed}
  [NAME, DATE]
\end{zed}

\begin{schema}{BirthdayBook}
  known: \power NAME \\
  birthday: NAME \pfun DATE
\where
  known = \dom  birthday
\end{schema}

\begin{theorem}{BirthdayBookPredicate}
\vdash? \forall BirthdayBook @ known = \dom birthday
\end{theorem}

\begin{zproof}[BirthdayBookPredicate]
reduce;
\end{zproof}

\begin{schema}{InitBirthdayBook}
  BirthdayBook~'
\where
  birthday' = \emptyset\\
  known' = \emptyset
\end{schema}

\begin{theorem}{InitIsOK}
\vdash? \exists  BirthdayBook~' @ InitBirthdayBook
\end{theorem}

\begin{zproof}[InitIsOK]
prove by reduce;
\end{zproof}

\begin{schema}{AddBirthday}
  \Delta BirthdayBook\\
  name?: NAME\\
  date?: DATE
\where
  name? \notin  known\\
  birthday' = birthday \cup  \{(name? \mapsto  date?)\}
\end{schema}

\begin{theorem}{AddBirthdayIsHonest}
\vdash? \forall  BirthdayBook; name?: NAME; date?: DATE @ name? \notin  known \implies  \pre  AddBirthday
\end{theorem}

\begin{zproof}[AddBirthdayIsHonest]
prove by reduce;
\end{zproof}

\begin{theorem}{knownAddBirthday}
\vdash? \forall AddBirthday @ known' = known \cup  \{name?\}
\end{theorem}

\begin{zproof}[knownAddBirthday]
prove by reduce;
\end{zproof}

\begin{schema}{FindBirthday}
  \Xi BirthdayBook\\
  name?: NAME\\
  date!: DATE
\where
  name? \in  known\\
  date! = birthday(name?)
\end{schema}

\begin{zproof}[FindBirthday\$domainCheck]
simplify;
\end{zproof}

\begin{zed}
AddThenFind == AddBirthday \semi  FindBirthday
\end{zed}

\begin{theorem}{AddFind}
\vdash? \forall AddThenFind @ date! = date?
\end{theorem}

\begin{zproof}[AddFind]
prove by reduce;
\end{zproof}

\begin{schema}{Remind1}
  \Xi BirthdayBook\\
  today?: DATE\\
  cards!: \power  NAME
\where
  cards! = \{  n: NAME | birthday~(n) = today? \}
\end{schema}

\begin{zproof}[Remind1\$domainCheck]
prove;
\end{zproof}

Here we had a problem before: the newline followed by "to" was being wrongly
captured by the lexer because of complex apply commands. We just changed the
lexing rule to disallow newlines before TO
\begin{schema}{Remind}
  \Xi BirthdayBook \\
  today?: DATE\\
  cards!: \power  NAME
\where
  cards! = \{  n: known | birthday(n) = today? \}
\end{schema}

\begin{zproof}[Remind\$domainCheck]
prove;
\end{zproof}

\begin{zed}
REPORT ::= ok | already\_known | not\_known
\end{zed}

\begin{schema}{Success}
  result!: REPORT
\where
  result! = ok
\end{schema}

\begin{schema}{AlreadyKnown}
  \Xi BirthdayBook\\
  name?: NAME\\
  result!: REPORT
\where
  name? \in  known\\
  result! = already\_known
\end{schema}

\begin{zed}
RAddBirthday == AddBirthday \land  Success \lor  AlreadyKnown
\end{zed}

\begin{theorem}{RAddBirthdayIsTotal}
\vdash? \forall  BirthdayBook; name?: NAME; date?: DATE @ \pre  RAddBirthday
\end{theorem}

\begin{zproof}[RAddBirthdayIsTotal]
prove by reduce;
split name? \in  known;
prove;
\end{zproof}

\begin{axdef}
  name: NAME\\
  VALUE: \nat
\where
  VALUE > 0
\end{axdef}

\begin{theorem}{tAddBirthdayPRE}
\vdash? \forall  BirthdayBook; name?: NAME; date?: DATE @ \pre  AddBirthday
\end{theorem}

\begin{zproof}[tAddBirthdayPRE]
invoke AddBirthday;
invoke \Delta BirthdayBook;
rewrite;
invoke BirthdayBook;
simplify;
cases;
next;
next;
next;
conjunctive;
simplify;
disjunctive;
simplify;
cases;
apply domInPower to predicate \dom  (birthday \cup  \{(name?, date?)\}) \in  \power  NAME;
rewrite;
next;
rewrite;
prove by reduce;
equality substitute \dom  birthday;
\end{zproof}

\begin{theorem}{myThm}
\vdash? \forall  x: \nat  @ x \geq  0
\end{theorem}

\begin{theorem}{test}
\vdash? \forall BirthdayBook @ true
\end{theorem}

\begin{zproof}[test]
invoke BirthdayBook;
\end{zproof}

