\begin{zsection}
   \SECTION zevesopt \parents standard\_toolkit
\end{zsection}

The "name on the thrid" parameter *must* not be the same as the unicode markup for
\verb'\after. That is, it cannot be just *after*. Otherwise, it gets confused later
on for the domain check and other theorems.
%
%%Zinword \after after
%%Zinword \before before
%%Zinword \middle middle
%%Zinword \zevesgen zevesgen
%%Zinword \zevesgenother zevesgenother
Zinword \zevesthmgen zevesthmgen
%%Zinword \zevesrel zevesrel
%%Zinword \zevesthmrel zevesthmrel

No need for ZEves syndef. Instead, define them as usual CZT / StdZ op templates.
The CZT2ZEves printer figures out what kind of ZEves optemp to use, and raise an
error in case they are not feasible. For instance, Z/EVES does not allow list ops
(e.g., seq display). And also, ZEves does not take assoc into account and precedences
are transformed (e.g., ZEves only have 1..6).
syndef{\after}{infun6}{"afterOp"}
syndef{\before}{infun6}{"beforeOp"}
syndef{\middle}{infun6}{"middleOp"}
syndef{\zevesgen}{ingen}{"zevesgen"}
syndef{\zevesrel}{inrel}{"zevesrel"}
syndef{\zevesthmrel}{inrel}{"zevesthmrel"}

\begin{zed}
\function 60 \leftassoc (\_ \after \_)
\also
\function 60 \leftassoc (\_ \before \_)
\also
\function 50 \leftassoc (\_ \middle \_)
\also
\generic 10 \leftassoc (\_ \zevesgen \_)
%\also
%\generic 10 \leftassoc (\_ \zevesthmgen \_)
\also
\generic 10 \leftassoc (\_ \zevesgenother\_)
\also
\relation  (\_ \zevesrel \_)
\also
\relation  (\_ \zevesthmrel \_)
\end{zed}

Example of an operator within an axdef
\begin{gendef}[X]
   \_ \after \_ : \seq~X \cross \nat \fun \seq~X
\where
   \Label{disabled rule dAfter}
   \forall s: \seq~X; n: \nat @ s \after n = \emptyset
\end{gendef}

The name here is the Unicode name of the operator name. That is, "after" because \after is "after"
\begin{zproof}[after\$domainCheck]
prove by reduce;
\end{zproof}

Example of operator within a schema both as declaration (E.g., include the _)
and as an application (e.g., don't include the _ in (a \before b)).
\begin{schema}{S}
   \_ \before \_ : \nat \cross \nat \fun \nat
\where
  \forall a, b: \nat @ a \before b > b
\end{schema}

\begin{zproof}[S\$domainCheck]
prove by reduce;
\end{zproof}

We are deliberately not accepting them in theorems like
\begin{theorem}{lOpInThm}
  \forall \_ \middle \_ : \nat \cross \nat \fun \nat @ 0 \middle 0 = 0
\end{theorem}

\begin{zed}
   X \zevesgen Y == \{ f: X \fun Y | f \neq \emptyset \}
\end{zed}

\begin{zed}
   \_ \zevesgenother \_ == (\nat \cross \nat) \rel \nat
\end{zed}

%begin{theorem}{lOpGenInThm}[X,Y]
%   \forall f: X \zevesthmgen Y @ f \neq \emptyset
%end{theorem}

\begin{gendef}[X, Y]
   \_ \zevesrel \_ : X \rel Y
\where
   \forall x: X ; y: Y @ (x \zevesrel y)
\end{gendef}

\begin{theorem}{lOpRelInThm}[X,Y]
   \forall \_ \zevesthmrel \_ : X \rel Y; x: X; y: Y @ x \zevesthmrel y
\end{theorem}
