\ai4fmignore{
\begin{zsection}
  \SECTION ch28 \parents ch27
\end{zsection}
CHANGES:
* missing hard spaces for DecorExpr
* \bot -> mondexError
* status -> statuS
}
\chapter[\Betw\ to \Conc\ lemmas]{Lemmas for the \Betw\ to \Conc\ correctness proofs} \label{ch28} %

\section{Specialising the proof rules}\label{ch28.bc-ignore-spec}
%-------------------

For each concrete operation $COp$ and corresponding between
operation $BOp$ we have to show

\begin{gzed}
%%\forall
  Rbc ; COp
%%@
  \shows \exists BetweenWorld' @ Rbc' \land BOp
\end{gzed}
Many operations are defined as the disjunction of other operations.
A $COp$ will have the same branches as a corresponding $BOp$: a
$CIgnore$ branch, and either a $CAbort$ or $COpOkay$ branch, or
both. We split the proof obligation into $CIgnore$, $CAbort$ and
$COpOkay$ branches, as we did in section~\ref{ch14.between-special}.
This gives some or all of the following proof requirements,
depending on which branches are in $COp$:
%
\begin{gzed}
%%(\forall
  Rbc ; CIgnore
%%@
  \shows \exists BetweenWorld' @ Rbc' \land Ignore
%%)\land
  \also %
%%(\forall
  Rbc ; CAbort
%% @
  \shows \exists BetweenWorld' @ Rbc' \land Abort
%%) \land
  \also %
%%(\forall
  Rbc ;  COpOkay
%%@
  \shows \exists BetweenWorld'  @ Rbc' \land BOpOkay
%%)
\end{gzed}
%
The correctness of the $CIgnore$ branch is dealt with below in
section \ref{ch28.bc-ignore-ignore}.  We then develop the
correctness proof for the $CAbort$ and $COpOkay$ branches, and
introduce a lemma applicable to certain operations.  Following this,
we present the proof of correctness of two common branches ---
$CIncrease$ and $CAbort$.

\section{Correctness of $CIgnore$}\label{ch28.bc-ignore-ignore}
%-------------------

The correctness of the $CIgnore$ branch follows trivially by
choosing

\begin{LNewThm}[\Betw\ to \Conc\ $CIgnore$ correct]
\begin{theorem}{tBetwConcCIgnoreCorrect}
    \forall Rbc ; CIgnore @ \exists BetweenWorld~' @ Rbc~' \land Ignore
\end{theorem}~\end{LNewThm}

Like $PhiBOpBranchesSig$
(see~\ref{ch8.betw.promotionschema.branches}), but with $ConPurse'$
from $AbortPurseOkay$.
%
\begin{LNewSDef}[$Abort \sqsubseteq CAbort$ correctness signature]
\begin{zed}
    BetwCAbortCorrectSig ~~\defs~~ PhiBOpBranchesSig \land AbortPurseOkay
\end{zed}~\end{LNewSDef}

\begin{LNewLemma}[$Abort \sqsubseteq CAbort$ correct $AuxWorld$ type]
\begin{theorem}{lAuxWorldCAbortType}
    \forall BetwCAbortCorrectSig @ AuxWorld[conAuthPurse := conAuthPurse \oplus \\
        \t1 \{~(name?, \theta ConPurse'[statuS' := eaFrom])~\}, ether := ether \cup \{~mondexError~\}]
\end{theorem}~\end{LNewLemma}

\begin{LNewLemma}[$Abort \sqsubseteq CAbort$ correct $BetweenWorld$ type]
\begin{theorem}{lBetweenWorldCAbortType}
    \forall BetwCAbortCorrectSig @ BetweenWorld[conAuthPurse := conAuthPurse \oplus \\
        \t1 \{~(name?, \theta ConPurse'[statuS' := eaFrom])~\}, ether := ether \cup \{~mondexError~\}]
\end{theorem}~\end{LNewLemma}

\begin{LNewThm}[\Betw\ \Conc\ correctness for $Abort$]
\begin{theorem}{tBetwConcCAbortCorrect}
    \forall Rbc ; CAbort @ \exists BetweenWorld~' @ Rbc~' \land Abort
\end{theorem}~\end{LNewThm}


\section{Correctness of a branch of the operation}\label{ch28.bc-ignore-okay}

Like in Section~\ref{ch14.between-special}, in here we need to
consider each $BOpOkay$ branch separately. The formal declaration of
these branches were given in
Sections~\ref{ch8.specproof-betw.promotedalone}
and~\ref{ch8.specproof-betw.worldlevelalone}.

As occurred in Chapter~\ref{ch14}, the proof strategy described in
Mondex uses meta-theorems to generalise the proof. For
mechanisation, we need to state those theorems explicitly.

\subsection{Theorems for each branch from $BOpOkay$ operation}

As $CIncrease$ and $CAbort$ are originally dealt with separately in
their own sections, we just add here the remaining cases.

\begin{LNewThm}[\Betw\ to \Conc\ $CStartFrom$ correct]
\begin{theorem}{tBetwConcCStartFromCorrect}
    \forall Rbc ; CStartFrom @ \exists BetweenWorld~' @ Rbc~' \land StartFrom %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CStartTo$ correct]
\begin{theorem}{tBetwConcCStartToCorrect}
    \forall Rbc ; CStartTo @ \exists BetweenWorld~' @ Rbc~' \land StartTo %
\end{theorem}~\end{LNewThm}

\begin{LNewSDef}[$Req \sqsubseteq CReq$ signature]
\begin{zed}
    BetwCReqCorrectSig ~~\defs~~ PhiBOpBranchesSig \land ReqPurseOkay
\end{zed}~\end{LNewSDef}

\begin{theorem}{lAuxWorldCReqType}
    \forall BetwCReqCorrectSig @ AuxWorld[conAuthPurse := conAuthPurse \oplus \\
        \t1 \{~(name?, \theta ConPurse'[statuS' := epa])~\}, ether := ether \cup \{~m!~\}]
\end{theorem}

\begin{theorem}{lBetweenWorldCReqType}
    \forall BetwCReqCorrectSig @ BetweenWorld[conAuthPurse := conAuthPurse \oplus \\
        \t1 \{~(name?, \theta ConPurse'[statuS' := epa])~\}, ether := ether \cup \{~m!~\}]
\end{theorem}

\begin{LNewThm}[\Betw\ to \Conc\ $CReq$ correct]
\begin{theorem}{tBetwConcCReqCorrect}
    \forall Rbc ; CReq @ \exists BetweenWorld~' @ Rbc~' \land Req %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CVal$ correct]
\begin{theorem}{tBetwConcCValCorrect}
    \forall Rbc ; CVal @ \exists BetweenWorld~' @ Rbc~' \land Val %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CAck$ correct]
\begin{theorem}{tBetwConcCAckCorrect}
    \forall Rbc ; CAck @ \exists BetweenWorld~' @ Rbc~' \land Ack %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CReadExceptionLog$ correct]
\begin{theorem}{tBetwConcCReadExceptionLogCorrect}
    \forall Rbc ; CReadExceptionLog @ \exists BetweenWorld~' @ Rbc~' \land ReadExceptionLog %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CClearExceptionLog$ correct]
\begin{theorem}{tBetwConcCClearExceptionLogCorrect}
    \forall Rbc ; CClearExceptionLog @ \exists BetweenWorld~' @ Rbc~' \land ClearExceptionLog %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CAuthoriseExLogClear$ correct]
\begin{theorem}{tBetwConcCAuthoriseExLogClearCorrect}
    \forall Rbc ; CAuthoriseExLogClear @ \exists BetweenWorld~' @ Rbc~' \land AuthoriseExLogClear %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CArchive$ correct]
\begin{theorem}{tBetwConcCArchiveCorrect}
    \forall Rbc ; CArchive @ \exists BetweenWorld~' @ Rbc~' \land CArchive %
\end{theorem}~\end{LNewThm}

\begin{LNewThm}[\Betw\ to \Conc\ $CIncrease$ correct]
\begin{theorem}{tBetwConcCIncreaseCorrect}
    \forall Rbc ; CIncrease @ \exists BetweenWorld~' @ Rbc~' \land Increase %
\end{theorem}~\end{LNewThm}



\begin{LNewThm}[\Betw\ to \Conc\ branching is correct]
\begin{theorem}{tBetwConcBranchingCorrect}
    (Rbc \land CIncrease \implies (\exists BetweenWorld~' @ Rbc~' \land Increase))\\ %
    \iff \\
    (
     (Rbc \land CIgnore \implies (\exists BetweenWorld~' @ Rbc~' \land Ignore)) \land \\ %
     (Rbc \land CAbort \implies (\exists BetweenWorld~' @ Rbc~' \land Abort)) \land \\ %
     (Rbc \land (~ \exists \Delta ConPurse @ PhiCOp \land IncreasePurseOkay ~) \\ %
        \t1 \implies (\exists BetweenWorld~' @ Rbc~' \land IncreaseOkay)))  %
\end{theorem}~\end{LNewThm}

\subsection{Choosing $BetweenWorld'$}\label{ch28.bc-choice}

TODO: CONTINUE TEXT FROM HERE

In choosing $BetweenWorld'$, we base our choice of the
$conAuthPurse'$ and $archive'$ components on $Rbc'$, and our choice
of the $ether'$ component on $BOpOkay'$.

We have $conAuthPurse_0'$ and $archive_0'$ in the hypothesis, and we
use this to provide the value for $conAuthPurse'$ and $archive'$,
respectively (this satisfies the constraint on $conAuthPurse'$ and
$archive'$ in $Rbc'$).
\begin{gzed}
%% \forall Rbc' @
  conAuthPurse' = conAuthPurse_0'
%%\land
  \\ %
  archive' = archive_0'
\end{gzed}
$m!$ and $ether$ are declared in the hypothesis, and $ether'$ can be
constructed deterministically from these (note that the following
construction satisfies the relevant constraint in $BOpOkay$ ---
either in $PhiBOp$ or explicitly as in $Archive$).
\begin{gzed}
%% \forall BOp @
  ether' = ether \cup \{ m! \}
\end{gzed}
We need to show that the chosen $BetweenWorld'$ and $m!$ satisfy
each of the conjuncts in the consequent (retrieve $Rbc'$ and
operation $BOpOkay$).

We also need to show that this choice is indeed an after
$BetweenWorld'$ (that it satisfies the constraints on $BetweenWorld$
specified in section~\ref{ch5.betweenworld}).
%............
\subsection{Case $BOpOkay$}
\label{bc-bop}
%............
From the choice of $ether'$ above, the relevant constraint on
$ether'$ in $BOpOkay$ is satisfied by construction.

At most one purse changes in $COpOkay$.  Let us call this new purse
value $p$.  This gives
\begin{argue}
  conAuthPurse_0' = conAuthPurse_0 \oplus \{ p \}
  \also %
  conAuthPurse_0' = conAuthPurse \oplus \{ p \} & $Rbc$
  \also %
  conAuthPurse' = conAuthPurse \oplus \{ p \} & choice of
  $conAuthPurse'$
\end{argue}
This satisfies the constraint on $conAuthPurse'$ in $BOpOkay$ (where
at most one purse changes in an identical manner to $COpOkay$).

$archive'$ is a function of $archive$ and $m!$, defined in
$BOpOkay$. Call this function $f$:
\begin{gaxdef}
  f: Logbook \cross MESSAGE \fun Logbook
\end{gaxdef}
Because $COpOkay$ is defined in an analogous way, $f$ also relates
$archive_0'$ to $archive_0$ and $m!$.

From the hypothesis we have $COpOkay$ and $Rbc$, and with our choice
of $archive'$ we have, respectively
\begin{gzed}
%%\forall COp; BOp @
  archive_0' = f(archive_0, m!)
  \\ %
  \land archive = archive_0
  \\ %
  \land archive' = archive_0'
\end{gzed}
Substituting the latter two equations into the first gives the
predicate in $BOpOkay$.

Thus, the $BOpOkay$ constraints on all the components of our chosen
$BetweenWorld'$ are satisfied under the correctness hypothesis and
choice of $BetweenWorld'$.

$\Box$ \ref{bc-bop}
%....................
\subsection{Case $Rbc'$}
\label{bc-rbc-dash}
%....................
Both the $conAuthPurse'$ and $archive'$ components of
$BetweenWorld'$ satisfy $Rbc'$ from the choice of $BetweenWorld'$.

All $COpOkay$ operations constrain $ether'$ as
\begin{gzed}
%% \forall COp @
  ether_0' \subseteq ether_0 \cup \{ m! \}
\end{gzed}
either through $PhiCOp$, or explicitly in $CArchive$.  Hence for
$ether'$ we have
\begin{argue}
  ether'
  \also %
  \t1 = ether \cup \{ m! \} & choice of $ether'$
  \also %
  \t1 \supseteq ether_0 \cup \{ m! \} & $Rbc$
  \also %
  \t1 \supseteq ether_0' & $COpOkay$
\end{argue}
This satisfies the constraint on $ether'$ in $Rbc'$.
%............
\subsection{Case `obey constraints'}
\label{bc-cons}
%............
We know from the hypothesis that the before $BetweenWorld$ satisfies
the constraints, so we need check only that the chosen message $m!$,
and any change of purse state during the operation, maintains this
constraint.

%\begin{lemma}[constraint]\index{lemma `constraint'}
  {\rm If an operation obeys the following properties, then it
    preserves the $BetweenWorld$ constraints:
    \begin{itemize}
    \item it does not change purse statuS or current transaction
      details ($pdAuth$)
    \item it does not change $allLogs$
    \item it does not change the payment detail messages, exception
      log read messages or exception log clear messages in the $ether$
      (either by not emitting such a message, or by emitting an
      already existing message)
    \item no sequence number decreases (all concrete operations have
      the property, so it is automatically satisfied)
    \end{itemize}
    }
%\end{lemma}
{\bf Proof:}

\noindent The $BetweenWorld$ constraints refer only to certain
$ether$ messages, namely,
\[
req,\ val,\ ack,\ exceptionLogResult,\ \mbox{and}\ exceptionLogClear %
\]
and relate their presence or absence to purse
status ($statuS$, $pdAuth$ and $nextSeqNo$) and $allLogs$. From the
hypothesis we can invoke lemma `logs unchanged'
(section~\ref{lemma-logs-unchanged}) to say that, as $allLogs$ does
not change, not does $alLogs$.  So operations that do not change the
purse status, do not change $allLogs$, and do not emit any relevant
new messages, will automatically preserve the constraints.

$\Box$ \ref{bc-cons}

\noindent Even when lemma `constraint' does not apply, we know from
the form of the operation that at most one purse changes, and one
message is emitted.  As at most one purse changes, the proof that
the $BetweenWorld$ constraints are preserved need refer only to this
purse; the constraints hold on the other purses before the operation
by hypothesis, and so they hold afterward, too.
%............
\subsection{Summary of $ConOkay$ proof obligation}
\label{bc-summary}
%............
For each operation, we have to show that either lemma `constraint'
holds or that the choice of $BetweenWorld'$ obeys the constraints
(see section~\ref{ch5.betweenworld}).
%-----------
\section{Correctness of $CIncrease$}
\label{bc-increase}
%-----------
$CIncrease$ does not change $statuS$ or $pdAuth$, does not log, and
no relevant message is emitted to the $ether$, so lemma `constraint'
(section \ref{lemma-constraint}) is applicable.

$\Box$ \ref{bc-increase}
%---------------------
\section{Correctness of $CAbort$}
\label{bc-abort}
%---------------------
Lemma `constraint' is not applicable, because $CAbort$ moves one
purse into $eaFrom$, and it may not have been in this state before,
and it may log a pending transaction.  Therefore we have to show
that our chosen $BetweenWorld'$ obeys the constraints.

One $mondexError$ message is emitted, and (possibly) one log is recorded.

\renewcommand{\labelenumi}{B--\theenumi}
\begin{enumerate}
\item $req \implies$ authentic $to$ purse.  No new $req$ messages.
\item No future $req$s.  No new $req$ messages.
\item No future $val$s.  No new $val$ messages.
\item No future $ack$s.  No new $ack$ messages.
\item No future $from$ logs.  The purse moves into $eaFrom$, possibly
  logging a transaction, and possibly increasing $nextSeqNo$.  This
  does not invalidate this constraint for any previous logs.  To
  create a new $from$ log, the purse would have had to have been in
  $epa$ (from {\it LogIfNecessary}).  Hence, using $ConPurse$
  constraint P--ref[p-epr], we have
  \[
  pdAuth.fromSeqNo < nextSeqNo
  \]
  From $AbortPurse$,  we also have
  \[
  nextSeqNo \leq nextSeqNo'
  \]
  This gives
  \[
  pdAuth.fromSeqNo < nextSeqNo'
  \]
  The $pdAuth$ is logged when the pre-state purse is in $epa$, and
  thus the new log obeys the constraint.

\item No future $to$ logs.  The purse moves into $eaFrom$, possibly
  logging a transaction, and possibly increasing $nextSeqNo$.  This
  does not invalidate this constraint for any previous logs.  To
  create a new $to$ log, the purse would have had to have been in
  $epv$ (from {\it LogIfNecessary}); hence, using $ConPurse$
  constraint P--ref[p-epr-name], we have
  \[
  pdAuth.toSeqNo < nextSeqNo
  \]
  From $AbortPurse$,  we also have
  \[
  nextSeqNo \leq nextSeqNo'
  \]
  This gives
  \[
  pdAuth.toSeqNo < nextSeqNo'
  \]
  The $pdAuth$ is logged when the pre-state purse is in $epv$, and
  thus the new log obeys the constraint.

\item $from$ in $\{epr,epa\}$, so no future $from$ logs.  The purse
  moves into $eaFrom$, so no new purses in $epr$ or $epa$.
\item $to$ in $\{epv,eaTo\}$, so no future $to$ logs.  The purse
  moves into $eaFrom$, so no new purses in $epv$ or $eaTo$.

\item $epr \implies \lnot val \land \lnot ack$.  The purse moves into
  $eaFrom$, and so does not move into $epr$.

\item $req \land \lnot ack \iff toInEpv \lor toLogged$.
  \begin{itemize}
  \item case $\implies$:

    No new $req$ messages; no $ack$ messages removed from the $ether$.

    The purse may have moved out of $epv$, but in such a case
    $LogIfNecessary$ says that it logs, hence re-establishing the
    condition.

  \item case $\Leftarrow$:

    No purses newly in $epv$.

    There might be a new $to$ log, in which case we must show there
    was a $req$, but no $ack$ before.  A $to$ log can be made only by
    a purse moving out of $epv$.  Then the $BetweenWorld$ constraint
    B--ref[b-req-no-ack], on $toInEpv$, before the operation gives us
    the required $req$ and lack of $ack$.
  \end{itemize}
\item $epv \land val \implies fromInEpa \lor fromLogged$.  No purses
  newly in $epv$; no new $val$ messages.

  The purse may have moved out of $epa$.  But in such a case
  $LogIfNecessary$ says that it logs, hence re-establishing the
  condition.
\item $fromInEpa \lor fromLogged \implies req$.  No purses newly in
  $epa$.

  There might be a new $from$ log, in which case we must show there
  was a $req$ before.  A $from$ log can be made only by a purse moving
  out of $epa$.  Then the $BetweenWorld$ constraint B--ref[b-is-req],
  on $fromInEpa$, before the operation gives us the required $req$.
\item $toLogged$ finite.  At most one $to$ log written, so finite
  before gives finite after.
\item $exceptionLogResult$s in $allLogs$.  No new exception log result
  messages.
\item Cleared logs archived.  No $exceptionLogClear$ messages are
  added, and the archive is unchanged.
\item $req$ for each log.  If there are no new logs, then the
  constraint holds from the pre-state.

  If a transaction exception is logged, then the purse status must
  have been either $epv$ or $epa$.  From constraints
  B--ref[b-req-no-ack] and B--ref[b-is-req], there was a $req$ in
  the pre-state $ether$ for the transaction which was logged.  This
  $req$ will also be in the post-state $ether$.
\end{enumerate}

$\Box$ \ref{bc-abort}

%---------------------
\section{Lemma `logs unchanged'}
\label{logs-unchanged}
%---------------------
%\begin{lemma}[logs unchanged]
  {\rm When the $archive$ and the individual purse logs do not change,
    and when no new $req$ messages are added to the $ether$, the set
    of $PayDetails$ representing all the logs does not change either.
    }
  \begin{gzed}
  BOpOkay | archive' = archive
  \\ %
  \t1 \land req \rres ether' = req \rres ether
  \\ %
  \t1 \land \forall n : \dom conAuthPurse @
  \\ %
  \t2 (conAuthPurse'~n).exLog =  (conAuthPurse~n).exLog
  \\ %
  \shows
  \\ %
  allLogs' = allLogs
  \\ %
  \land toLogged' = toLogged
  \\ %
  \land fromLogged' = fromLogged
  \end{gzed}
%\end{lemma}

\noindent{\bf Proof:}
\begin{argue}
  allLogs = archive
  \\ %
  \t2 {} \cup \{~ n:\dom conAuthPurse; ld:PayDetails |
  \\ %
  \t3 ld \in (conAuthPurse~n).exLog ~\}
  \\ %
  & defn
  \also %
  \t1 = archive'
  \\ %
  \t2 {} \cup \{~ n:\dom conAuthPurse'; ld:PayDetails |
  \\ %
  \t3 ld \in (conAuthPurse'~n).exLog ~\}
  \\ %
  & assumption and $PhiBOp$
  \also %
  \t1 = allLogs' & defn
  \also %
  allLogs =  \{~ n:\dom conAuthPurse; pd:PayDetails |
  \\ %
  \t2 n \mapsto pd \in allLogs \land req~pd \in ether ~ \}
  \\ %
  & defn

  \also %
  \t1 =  \{~ n:\dom conAuthPurse'; pd:PayDetails |
  \\ %
  \t2 n \mapsto pd \in allLogs' \land req~pd \in ether' ~ \}
  \\ %
  & assumption and  above
  \also %
  \t1 = allLogs'
  \\ %
  & defn
\end{argue}
The arguments for $toLogged$ and $fromLogged$ follow in exactly the
same way.

$\Box$ \ref{logs-unchanged}
%---------------------
\section{Lemma `abort forward': operations that first abort}
\label{ignore-abort-bc-op}
%---------------------
Some concrete operations are written as a composition of $Abort$ and
a simpler operation starting from $eaFrom$ ($StartFrom$, $StartTo$,
$ReadExceptionLog$, {\it ClearExceptionLog}, etc.).

%\begin{lemma}[abort forward]
  {\rm Where a \Conc\ operation is written as a composition of
    $CAbort$ and a simpler operation starting from $eaFrom$, and the
    corresponding \Betw\ operation is structured analogously, it is
    sufficient to prove that the simpler \Conc\ operation refines the
    corresponding \Betw\ operation.  }
  \begin{gzed}
  (CAbort \semi COpEafrom); Rbc;
  \\ %
  \t2 (\forall COpEafrom; Rbc @ \exists BetweenWorld' @ Rbc' \land
  BOpEafrom)
  \\ %
  \shows
  \\ %
  \exists BetweenWorld' @ Rbc' \land (Abort \semi BOpEafrom)
  \end{gzed}
%\end{lemma}

\noindent{\bf Proof} We have already proved in section
\ref{bc-abort} that $CAbort$ refines $Abort$.  Adding this to our
hypothesis, we get
\begin{gzed}
(CAbort \semi COpEafrom); Rbc;
\\ %
\t2 (\forall CAbort; Rbc @ \exists BetweenWorld' @ Rbc' \land
Abort);
\\ %
\t2 (\forall COpEafrom; Rbc @ \exists BetweenWorld' @ Rbc' \land
BOpEafrom)
\\ %
\shows
\\ %
\exists BetweenWorld' @ Rbc' \land (Abort \semi BOpEafrom)
\end{gzed}
The hypothesis is now in precisely the form required to use lemma
`compose forward', (section~\ref{lemma-compose-fwd}) and we do so to
prove the consequent.

$\Box$ \ref{ignore-abort-bc-op}


\newpage
\section{Summary}\label{ch28.summary}

\ldefsummary %
\lthmsummary %
\lthmaddeddefsummary %
\lthmaddedthmsummary %
\lzevessummary %
%\lpscriptsummary
