\begin{zsection}
   \SECTION refRulesEquiv \parents zeves\_toolkit
\end{zsection}

This file shows an equivalence lemma (last theorem below) for a simplified
version of the Z IO refinement rules as described in the University of York
report YCS-2002-347 by Iain Toyn, Susan Stepney, and Jim Woodcock.

\begin{schema}{AS}
%\zastate\\
   x: \power~\nat
\end{schema}
\begin{zed}
   FI ::= NatIn \ldata \nat \rdata
 \also
   FO ::= NatOut \ldata \nat \rdata
\end{zed}

\begin{schema}{CS}
%\zcstate\\
   y: \seq~\nat
\end{schema}

The next refinement tag is mistaken on purpose - just to test the parser for it.
\begin{schema}{AOper}
   \Delta AS\\
   ai? : \nat\\
   ao! : \nat
\where
   x' = x \cup \{ ai? \}
   \\
   ao! \in x
\end{schema}

\begin{schema}{COper}
%\zfsrefines AOper\\
   \Delta CS \\
   ci?: FI \\
   co!: FO
\where
   y' = y \cat \langle (NatIn\inv)~(ci?) \rangle \\
   (NatOut\inv)~(co!) \in \ran~y
\end{schema}

\begin{schema}{R}
%\zretrieve\\
   AS; CS
\where
   x = \ran~y
\end{schema}

\begin{schema}{RIn}
%\zretrievein\\
   ai?: \nat \\
   ci?: FI
\where
   ai? = (NatIn\inv)~(ci?)
\end{schema}

\begin{schema}{ROut}
%\zretrieveout\\
   ao!: \nat \\
   co!: FO
\where
   ao! = (NatOut\inv)~(co!)
\end{schema}

\begin{theorem}{BSCorrectnessEquiv0} 
  (\forall CS; ci?: FI  | (\forall AS; ai?: \nat | R \land RIn @ \pre AOper) @
  	(\forall CS~'; co!: FO; AS~'; ao!: \nat | COper \land R~' \land ROut @
  		(\exists AS; ai?: \nat @ R \land RIn \land AOper)))
\end{theorem}

\begin{theorem}{rule lNatInType}
   \forall x: FI @ (NatIn\inv)~x \in \num
\end{theorem}

\begin{theorem}{rule lNatInIsNat}
   \forall x: FI @ (NatIn\inv)~x \geq 0
\end{theorem}

\begin{theorem}{BSCorrectnessEquiv} 
  (\forall CS; ci?: FI  | (\forall AS; ai?: \nat | R \land RIn @ \pre AOper) @
  	(\forall CS~'; co!: FO; AS~'; ao!: \nat | COper \land R~' \land ROut @
  		(\exists AS; ai?: \nat @ R \land RIn \land AOper)))
  	\iff
  (\forall COper; ROut; R~' | (\forall AS; RIn | R @ \pre AOper) @ (\exists AS; R; RIn @ AOper))
\end{theorem}

\begin{zproof}[BSCorrectnessEquiv]
split (\forall CS; ci?: FI  | (\forall AS; ai?: \nat | R \land RIn @ \pre AOper) @
  	(\forall CS~'; co!: FO; AS~'; ao!: \nat | COper \land R~' \land ROut @
  		(\exists AS; ai?: \nat @ R \land RIn \land AOper)));
cases;
prove by reduce;
instantiate x == \ran~y, y\_\_0 == y;
prove by reduce;
instantiate ci\_\_1? == ci?, ai\_\_0? == (NatIn\inv)~(ci?);
prove by reduce;
next;
split (\forall COper; ROut; R~' | (\forall AS; RIn | R @ \pre AOper) @ (\exists AS; R; RIn @ AOper));
simplify;
prove by reduce;
next;
\end{zproof}
