\documentclass{article}
\usepackage{czt}

\begin{document}
%----------------------------------

\begin{zsection}	 \SECTION zstateref\_ref \parents~zstateref
\end{zsection}

VC ZSection Created at Thu Sep 08 16:45:22 BST 2011.


Added schema for feasibility VC signature of paragraph named AOp1Sig at unknown.
\begin{schema}{AOp1Sig}
i? : \nat \\
 x : \power \nat 
\where
 AState
\end{schema}


Added schema for feasibility VC signature of paragraph named AOp2Sig at unknown.
\begin{schema}{AOp2Sig}
x : \power \nat 
\where
 AState
\end{schema}


Added schema for feasibility VC signature of paragraph named ASFinSig at unknown.
\begin{schema}{ASFinSig}
x : \power \nat 
\where
 ASFin
\end{schema}


Added schema for feasibility VC signature of paragraph named ASInitSig at unknown.
\begin{schema}{ASInitSig}
x' : \power \nat 
\where
 ASInit
\end{schema}


Added schema for feasibility VC signature of paragraph named AStateSig at unknown.
\begin{schema}{AStateSig}
x : \power \nat 
\where
 AState
\end{schema}


Added schema for feasibility VC signature of paragraph named COp1Sig at unknown.
\begin{schema}{COp1Sig}
i? : \nat \\
 y : \seq \nat 
\where
 CState
\end{schema}


Added schema for feasibility VC signature of paragraph named COp2Sig at unknown.
\begin{schema}{COp2Sig}
y : \seq \nat 
\where
 CState
\end{schema}


Added schema for feasibility VC signature of paragraph named CSFinSig at unknown.
\begin{schema}{CSFinSig}
y : \seq \nat 
\where
 CSFin
\end{schema}


Added schema for feasibility VC signature of paragraph named CSInitSig at unknown.
\begin{schema}{CSInitSig}
y' : \seq \nat 
\where
 CSInit
\end{schema}


Added schema for feasibility VC signature of paragraph named CStateSig at unknown.
\begin{schema}{CStateSig}
y : \seq \nat 
\where
 CState
\end{schema}


Added schema for feasibility VC signature of paragraph named NothingAgainSig at unknown.
\begin{schema}{NothingAgainSig}
y : \seq \nat 
\where
 CState
\end{schema}


Added schema for feasibility VC signature of paragraph named NothingSig at unknown.
\begin{schema}{NothingSig}
x : \power \nat 
\where
 AState
\end{schema}


Added schema for feasibility VC signature of paragraph named RetrSig at unknown.
\begin{schema}{RetrSig}
x : \power \nat \\
 y : \seq \nat 
\where
 Retr
\end{schema}

VC3 for line 6 column 1 start 81 length 90 in "zstateref"
\begin{theorem}{AState\_ vc\_ ref}\vdash ? \exists AStateSig | true @ true
\end{theorem}

VC9 for line 27 column 1 start 329 length 94 in "zstateref"
\begin{theorem}{CState\_ vc\_ ref}\vdash ? \exists CStateSig | true @ true
\end{theorem}

VC11 for line 34 column 1 start 425 length 91 in "zstateref"
\begin{theorem}{CSInit\_ vc\_ ref}\vdash ? \exists CSInitSig | true @ true
\end{theorem}

VC13 for line 41 column 1 start 518 length 70 in "zstateref"
\begin{theorem}{CSFin\_ vc\_ ref}\vdash ? \exists CSFinSig | true @ true
\end{theorem}

VC15 for line 48 column 1 start 590 length 90 in "zstateref"
\begin{theorem}{Retr\_ vc\_ ref}\vdash ? \exists RetrSig | true @ true
\end{theorem}

VC17 for line 57 column 1 start 714 length 68 in "zstateref"
\begin{theorem}{Nothing\_ vc\_ ref}\vdash ? \forall NothingSig | true @ \pre Nothing
\end{theorem}

VC19 for line 63 column 1 start 784 length 73 in "zstateref"
\begin{theorem}{NothingAgain\_ vc\_ ref}\vdash ? \forall NothingAgainSig | true @ \pre NothingAgain
\end{theorem}

VC21 for line 69 column 1 start 859 length 94 in "zstateref"
\begin{theorem}{AOp1\_ vc\_ ref}\vdash ? \forall AOp1Sig | true @ \pre AOp1
\end{theorem}

VC23 for line 76 column 1 start 955 length 125 in "zstateref"
\begin{theorem}{COp1\_ vc\_ ref}\vdash ? \forall COp1Sig | true @ \pre COp1
\end{theorem}

VC25 for line 84 column 1 start 1082 length 84 in "zstateref"
\begin{theorem}{AOp2\_ vc\_ ref}\vdash ? \forall AOp2Sig | true @ \pre AOp2
\end{theorem}

VC27 for line 91 column 1 start 1168 length 106 in "zstateref"
\begin{theorem}{COp2\_ vc\_ ref}\vdash ? \forall COp2Sig | true @ \pre COp2
\end{theorem}

VC29 for line 6 column 1 start 81 length 90 in "zstateref"
\begin{theorem}{tAStateInitialisation\_ vc\_ ref}\vdash ? \exists AState~' | true @ ASInit
\end{theorem}

VC30 for line 6 column 1 start 81 length 90 in "zstateref"
\begin{theorem}{tAStateFinalisation\_ vc\_ ref}\vdash ? \exists AState | true @ ASFin
\end{theorem}

VC31 for line 27 column 1 start 329 length 94 in "zstateref"
\begin{theorem}{tFSInitialisationCState\_ vc\_ ref}\vdash ? \forall CSInit | true @ \exists Retr~' | true @ ASInit
\end{theorem}

VC32 for line 41 column 1 start 518 length 70 in "zstateref"
\begin{theorem}{tFSFinalisationCState\_ vc\_ ref}\vdash ? \forall CSFin ; Retr | true @ ASFin
\end{theorem}

VC33 for line 76 column 1 start 955 length 125 in "zstateref"
\begin{theorem}{tFSApplicabilityCOp1\_ vc\_ ref}\vdash ? \forall AOp1Sig ; COp1Sig ; Retr | \pre AOp1 @ \pre COp1
\end{theorem}

VC34 for line 76 column 1 start 955 length 125 in "zstateref"
\begin{theorem}{tFSCorrectnessCOp1\_ vc\_ ref}\vdash ? \forall AState ; COp1 ; Retr | \pre AOp1 @ \exists AState~' | Retr~' @ AOp1
\end{theorem}

VC35 for line 91 column 1 start 1168 length 106 in "zstateref"
\begin{theorem}{tFSApplicabilityCOp2\_ vc\_ ref}\vdash ? \forall AOp2Sig ; COp2Sig ; Retr | \pre AOp2 @ \pre COp2
\end{theorem}

VC36 for line 91 column 1 start 1168 length 106 in "zstateref"
\begin{theorem}{tFSCorrectnessCOp2\_ vc\_ ref}\vdash ? \forall AState ; COp2 ; Retr | \pre AOp2 @ \exists AState~' | Retr~' @ AOp2
\end{theorem}



 VC Z section "zstateref_ref" has $19$ interesting VCs (Total = 36; Simplified to $true$ = 17).



%----------------------------------
\end{document}