\documentclass{article}
\usepackage{oz}
\begin{document}

\begin{zsection}
\SECTION emptiness\_tests \parents standard\_toolkit
\end{zsection}

Test schemas with empty declaration parts.

\begin{zed}
  TRUE == [|true]
\end{zed}

\begin{schema}{FALSE}
\where
  false
\end{schema}


Test axdef with no declarations.

\begin{axdef}
\where
   (\exists x:\nat | x < 10 @ x*x = x)
\end{axdef}


Test empty axdef.

\begin{axdef}
\end{axdef}



\begin{zsection}
\SECTION predicate\_tests \parents standard\_toolkit
\end{zsection}

Test chains together with member predicate.

\begin{schema}{MemberChain}
\where
  1 \in \{ 1 \} \in \power \{1\}
\end{schema}



\begin{zsection}
\SECTION nested\_predicate\_tests \parents standard\_toolkit
\end{zsection}

This tests the parsing of nested predicates that contain newlines
instead of conjunctions.

\begin{axdef}
  a, b:\nat
\end{axdef}

This predicate part of this schema should parse into
two separate predicates, first the forall, then the $y<a$,
because the newline conjunction has lower priority than the forall.
\begin{schema}{PredTest1}
   x:\nat \\
   y:\nat
\where
   \forall a:\nat \\
   \t1     b:\nat | \\
   \t1     a < 3 \\
   \t1     b < 4 @ \\
   \t1   x < a \\
   y < a
\end{schema}

This one should parse the same as the above!
\begin{schema}{PredTest2}
   x:\nat \\
   y:\nat
\where
   (\forall a:\nat \\
   \t1      b:\nat | \\
   \t1     a < 3 \\
   \t1     b < 4 @ \\
   \t1   x < a \\
   \t1   y < b)
\end{schema}

This is how one has to write it to have a conjunction using newlines
INSIDE the forall.
\begin{schema}{PredTest3}
   x:\nat \\
   y:\nat
\where
   \forall a:\nat \\
   \t1      b:\nat | \\
   \t1     a < 3 \\
   \t1     b < 4 @ \\
   \t1   (x < a \\
   \t1   y < b)
\end{schema}

This should parse the same as the first one (but uses ;).
\begin{schema}{SemicolonTest1}
   x:\nat \\
   y:\nat
\where
   (\forall a:\nat; \\
   \t1      b:\nat | \\
   \t1     a < 3 ; \\
   \t1     b < 4 @ \\
   \t1   x < a ; \\
   \t1   y < b)
\end{schema}

This should parse the same as the first one (but with newlines before
the operators rather than after).
\begin{schema}{SemicolonTest2} \\
   x \\ : \\ \nat \\
   y \\ : \\ \nat \\
\where
   (\forall a:\nat \\
   \t1      ;b:\nat \\
   \t1   | a < 3 \\
   \t1     b < 4 \\
   \t1   @ x < a
   \t1   ; y < b)
\end{schema}


\begin{zsection}
\SECTION bugs \parents standard\_toolkit
\end{zsection}

\begin{zed}
  Foo == [x:\nat]~_1
\end{zed}

\end{document}
