\documentstyle[11pt,oz]{article}
\zedindent -1mm
\pagestyle{empty}
\raggedbottom
\begin{document}
\begin{class}{DMS}[X]
\also
\begin{state}
  nodes: \finset X  \\
  ddo:  X  \rel  X  \\
  \Delta\\
  tc:  X  \rel  X  
\where
  \dom (ddo) \subseteq nodes \land \\ 
  tc = ddo\plus \land \\ 
  \lnot (\exists x:  X @ (x, x) \in tc)
\end{state}
\begin{init}
  nodes' = \emptyset \\ 
  ddo' = \emptyset
\end{init}
\begin{schema}{NoNodes}
  result!:  \bool 
\where
  result! = True \iff nodes = \emptyset \\ 
\end{schema}
\begin{schema}{IsNode}
  x?:  X  \\
  result!:  \bool  
\where
  result! = True \iff x? \in nodes
\end{schema}
\begin{schema}{AddNode}
  \Delta (nodes)\\
  x?:  X  
\where
  x? \notin nodes \\ 
  nodes' = nodes \cup \{x?\}
\end{schema}
\begin{schema}{RemoveNode}
  \Delta (nodes, ddo, tc)\\
  x?:  X  
\where
  x? \in nodes \setminus  \ran ddo \\
  nodes' = nodes \setminus \{x?\} \land \\ 
  ddo' = \{x?\} \ndres ddo
\end{schema}
\begin{schema}{NoDependencies}
  result!:  \bool  
\where
  result! = True \iff ddo = \emptyset
\end{schema}
\begin{schema}{DependedUpon}
  x?:  X  \\
  result!:  \bool  
\where
  x? \in nodes\\
  result! = True \iff x? \in  \ran ddo
\end{schema}
\begin{schema}{IsDependency}
  x?, y?:  X  \\
  result!:  \bool  
\where
  \{x?, y?\} \subseteq nodes \\ 
  result! = True \iff (x?, y?) \in ddo 
\end{schema}
\begin{schema}{CanAdd}
  x?, y?:  X  \\
  result!:  \bool  
\where
  \{x?, y?\} \subseteq nodes \\ 
  result! = True \iff ((y?, x?) \notin tc \land x? \neq y?)
\end{schema}
\begin{schema}{AddDependency}
  \Delta (ddo, tc)\\
  x?, y?:  X  
\where
  \{x?, y?\} \subseteq nodes \\
  x? \neq y? \\
  (y?, x?) \notin tc \\ 
  ddo' = ddo \cup \{(x?, y?)\}
\end{schema}
\begin{schema}{RemoveDependency}
  \Delta (ddo, tc)\\
  x?, y?:  X  
\where
  (x?, y?) \in ddo \\ 
  ddo' = ddo \setminus \{(x?, y?)\}
\end{schema}
\begin{schema}{Dependents}
  x?:  X  \\
  ns!:  \finset  X  
\where
  x? \in nodes \\ 
  ns! = \{n: nodes | (n, x?) \in tc\}
\end{schema}
\begin{schema}{Supporters}
  x?:  X  \\
  ns!: \finset  X  
\where
  x? \in nodes \\ 
  ns! = \{n: nodes | (x?, n) \in tc\}
\end{schema}
\begin{schema}{UltSupporters}
  x?:  X  \\
  ns!: \finset  X  
\where
  x? \in nodes \\ 
  ns! = \{n: nodes | (x?, n) \in tc \land n \notin  \dom ddo\}
\end{schema}
\begin{schema}{CandidateSupporters}
  x?:  X  \\
  cs!:  \finset  X 
\where
  x? \in nodes \\ 
  cs! = \{ n: nodes | (n, x?) \notin tc \land \\ n \neq x?\}
\end{schema}
\begin{schema}{SomeDirectDependent}
  x?:  X  \\
  n!:  X  
\where
  x? \in  \ran ddo \\ 
  (n!, x?) \in ddo
\end{schema} 
\end{class} 
\end{document}
