\documentclass{article}
\usepackage{czt}
\newcommand{\sexprUnfoldsTo}{\mathrel{=_{se}}}
\newcommand{\declListUnfoldsTo}{\mathrel{=_d}}
\newcommand{\unprefix}{\mathrel{unprefix}}
\newcommand{\schemaEquals}{\mathrel{=_S}}

\title{Schema Unfolding Rules in CZT}
\author{Mark Utting}
\begin{document}
\maketitle

\begin{zsection}
  \SECTION unfold \parents zpattern\_toolkit, standard\_toolkit
\end{zsection}

Declare the jokers used in these rules.

\begin{zedjoker}{DeclList} D, D1, D2, D3, D4, D5, D6, D7, D8, D9 \end{zedjoker}
\begin{zedjoker}{Pred} P, P1, P2, P3, P4, P5, P6, P7, P8, P9 \end{zedjoker}
\begin{zedjoker}{Expr} E, E1, E2, E3, E4, E5, E6, E7, E8, E9 \end{zedjoker}
\begin{zedjoker}{Name} v, v1, v2, v3, v4, v5, v6, v7, v8, v9 \end{zedjoker}

To clearly distinguish our rules for unfolding declaration lists and
schema expressions from other rules, we introduce two new infix
operators: $\sexprUnfoldsTo$ and $\declListUnfoldsTo$.  
We do not really need to define their semantics in order to use them within
rules, but to reassure readers, we define their semantics to be just
equality.  In fact, the intention is that the right hand argument
will be the unfolded and normalized form of the left hand argument.

\begin{verbatim}
%%Zinword \sexprUnfoldsTo sexprUnfoldsTo
%%Zinword \declListUnfoldsTo declListUnfoldsTo
\end{verbatim}

\begin{zed}
  \relation ( \_ \sexprUnfoldsTo \_ )
\end{zed}
\begin{zed}
  \relation ( \_ \declListUnfoldsTo \_ )
\end{zed}


\begin{gendef}[SCHEMA]
  \_ \sexprUnfoldsTo \_ : SCHEMA \rel SCHEMA \\
  \_ \declListUnfoldsTo \_ : SCHEMA \rel SCHEMA
\where
  \forall s1,s2:SCHEMA @ s1 \sexprUnfoldsTo s2 \iff s1=s2 \\
  \forall s1,s2:SCHEMA @ s1 \declListUnfoldsTo s2 \iff s1=s2 \\
\end{gendef}

\section{Rules for Unfolding Predicates}

\begin{zedrule}{SchemaPred}
  E \iff \theta E \in E
\end{zedrule}

%% Syntax error on the conclusion of these rules.
%\begin{zedrule}{implies}
%   (P1 \implies P2) \iff ((\lnot P1) \lor P2)
%\end{zedrule}

%\begin{zedrule}{iff}
%   (P1 \iff P2) \iff ((P1 \land P2) \lor ((\lnot P1) \land (\lnot P2))
%\end{zedrule}

\section{Unfolding Declaration Lists}

We use the $\declListUnfoldsTo$ operator for unfolding declaration
lists.   The left-hand side is always of the form $[DeclList|true]$
and the right-hand side (which is usually a generated output) is
a normalized schema with no duplicated names in the signature, and
all the non-basetype conditions in the predicate part.

The next few rules implement the $\declListUnfoldsTo$ operator,
which unfolds and normalizes declaration lists.

\begin{zedrule}{VarDecl1}
   \proviso E : \power E \\
   [D1 | true] \declListUnfoldsTo [D2 | P2]
\derives
   [v:E; D1 | true] \declListUnfoldsTo [v:E; D2 |  P2]
\end{zedrule}

\begin{zedrule}{VarDecl2}
   \proviso E : \power E2 \\
   [D1 | true] \declListUnfoldsTo [D2 | P2]
\derives
   [v:E; D1 | true] \declListUnfoldsTo [v:E2; D2 |  v \in E \land P2]
\end{zedrule}

\begin{zedrule}{ConstDecl}
   \proviso E : E2 \\
   [D1 | true] \declListUnfoldsTo [D2 | P2]
\derives
   [v==E; D1 | true] \declListUnfoldsTo [v:E2; D2 |  v = E \land P2]
\end{zedrule}

\begin{zedrule}{IncludeDecl}
   E \sexprUnfoldsTo [D1 | P1] \\
   [D | true] \declListUnfoldsTo [D2 | P2] \\
   \proviso [D1 | true] \land [D2 | true] : \power [D3]
\derives
   [E; D | true] \declListUnfoldsTo [D3 |  P1 \land P2]
\end{zedrule}

\begin{zedrule}{EmptyDeclList}
   [~ | true] \declListUnfoldsTo [~ | true]
\end{zedrule}


\section{Unfolding Schema Expressions}

This section defines the unfolding of schema expressions,
using the $se \sexprUnfoldsTo stext$ operator, where $se$
is a schema expression and $stext$ is the resulting normalized
schema construction ($[Decls|Preds]$, where the types in $Decls$
are only base types).

Top-level schema expressions are unfolded into schema construction
expressions, which can then be expanded into sets of bindings.
However, we have a special case for $\exists$, so that we can
put all the variables into the bound variable list of the
set comprehension, then create a binding using a subset of
those variables.  Putting all the variables at the same level
of scope gives the evaluation optimization algorithms more
freedom to reorder things later.

\begin{zedrule}{TopLevel}
  \proviso E : \power [D2] \\
  E  \sexprUnfoldsTo [D | P]
\derives
  E = [D | P]
\end{zedrule}

\begin{zedrule}{Theta}
  \proviso E : \power [D] \\
  \proviso E2 == \theta [D | true]
\derives
  \theta E = E2
\end{zedrule}

NOTE: we need thirteen versions of this next rule -- one for
each possible decoration.  The long-term solution is to add
support for Stroke-List jokers.
\begin{zedrule}{ThetaPrime}
  \proviso E : \power [D] \\
  \proviso E2 == \theta [D | true] '
\derives
  \theta E~' = E2
\end{zedrule}

\begin{zedrule}{SchemaConstruction1}
  [D1 | true] \declListUnfoldsTo [D2 | true]
\derives
  [D1 | P] \sexprUnfoldsTo [D2 | P]
\end{zedrule}

\begin{zedrule}{SchemaConstruction2}
  [D1 | true] \declListUnfoldsTo [D2 | P2]
\derives
  [D1 | P] \sexprUnfoldsTo [D2 | P2 \land P]
\end{zedrule}

\begin{zedrule}{SchemaRef}
  \proviso ?~ E == E2 \\
  E2 \sexprUnfoldsTo [D2 | P2]
\derives
  E \sexprUnfoldsTo [D2 | P2]
\end{zedrule}

This rule unfolds any remaining $\Delta$ schemas.
If the specification explicitly defined the $\Delta$ schema,
then the above SchemaRef rule would have unfolded it.
\begin{zedrule}{DeltaRef}
  \proviso v2 == \Delta \unprefix v \\
  [v2; v2~'] \sexprUnfoldsTo [D1|P1]
\derives
  v \sexprUnfoldsTo [D1|P1]
\end{zedrule}

This rule unfolds any remaining $\Xi$ schemas.
If the specification explicitly defined the $\Xi$ schema,
then the above SchemaRef rule would have unfolded it.
\begin{zedrule}{XiRef}
  \proviso v2 == \Xi \unprefix v \\
  [v2; v2~'] \sexprUnfoldsTo [D1|P1] \\
  \proviso [v2|true] : \power [D2]
\derives
  v \sexprUnfoldsTo [D1|P1 \land \theta [D2|true] = \theta [D2|true]~']
\end{zedrule}

NOTE: we need thirteen versions of this next rule -- one for
each possible decoration.  The long-term solution is to add
support for Stroke-List jokers.
\begin{zedrule}{SchemaPrime}
   E \sexprUnfoldsTo [D1 | P1] \\
   \proviso [D2|P2] == [D1|P1]~' \\
\derives
   E~' \sexprUnfoldsTo [D2 | P2]
\end{zedrule}

The type proviso in the ExistsSchema rule checks
that $D1$ and $D2$ are type compatible.
\begin{zedrule}{ExistsSchema}
   [D|P] \sexprUnfoldsTo [D1 | P1] \\
   E2 \sexprUnfoldsTo [D2 | P2] \\
   \proviso [D1 | true] \land [D2 | true] : \power [D3] \\
   \proviso [D4|true] == [D2|true] \schemaminus [D1|true]
\derives
   (\exists D | P @ E2) \sexprUnfoldsTo [D4 | (\exists D1 @ P1 \land P2)]
\end{zedrule}



\begin{zedrule}{SchemaNegation}
  E \sexprUnfoldsTo [D | P]
\derives
  (\lnot E) \sexprUnfoldsTo [D | \lnot P]
\end{zedrule}

\begin{zedrule}{SchemaConjunction}
  E1 \sexprUnfoldsTo [D1 | P1] \\
  E2 \sexprUnfoldsTo [D2 | P2] \\
 \proviso [D1 | true] \land [D2 | true] : \power [D3]
\derives
  (E1 \land E2) \sexprUnfoldsTo [D3 | P1 \land P2]
\end{zedrule}

\begin{zedrule}{SchemaDisjunction}
  E1 \sexprUnfoldsTo [D1 | P1] \\
  E2 \sexprUnfoldsTo [D2 | P2] \\
 \proviso [D1 | true] \land [D2 | true] : \power [D3]
\derives
  (E1 \lor E2) \sexprUnfoldsTo [D3 | P1 \lor P2]
\end{zedrule}

\begin{zedrule}{SchemaImplication}
  E1 \sexprUnfoldsTo [D1 | P1] \\
  E2 \sexprUnfoldsTo [D2 | P2] \\
 \proviso [D1 | true] \land [D2 | true] : \power [D3]
\derives
  (E1 \implies E2) \sexprUnfoldsTo [D3 | P1 \implies P2]
\end{zedrule}

\begin{zedrule}{SchemaEquivalence}
  E1 \sexprUnfoldsTo [D1 | P1] \\
  E2 \sexprUnfoldsTo [D2 | P2] \\
 \proviso [D1 | true] \land [D2 | true] : \power [D3]
\derives
  (E1 \iff E2) \sexprUnfoldsTo [D3 | P1 \iff P2]
\end{zedrule}

\begin{zedrule}{SchemaProjection}
  E1 \sexprUnfoldsTo [D1 | P1] \\
  E2 \sexprUnfoldsTo [D2 | P2] \\
  \proviso [D1 | true] \land [D2 | true] : \power [D3] \\
  \proviso [D4 | true] == [D3 | true] \schemaminus [D2 | true]
\derives
  E1 \project E2 \sexprUnfoldsTo [D2 | (\exists D4 @ P1 \land P2)]
\end{zedrule}

postnames : names that end with ! or '
prenames: names that do not end with ! or '
(that is, names that end with ?, or numeric decoration, or that 
don't have decorations).

\begin{zedrule}{SchemaPrecondition}
  E \sexprUnfoldsTo [D4 | P2] \\
  \proviso [D2 | true] == postnames [D4 | true] \\
  \proviso [D3 | true] == [D4 | true] \schemaminus [D2 | true] \\
\derives
  \pre E \sexprUnfoldsTo [D3 | (\exists D2 @ P2)]
\end{zedrule}

TODO: schema hiding, schema piping and sequential composition,
schema forall, schema exists unique.

This rule applies the above schema unfolding rules to unfolding
the bound variables of quantifiers, because the rewrite tactics
transform schema text using the $\schemaEquals$ relation.
\begin{zedrule}{Quantifiers}
   [D|P] \sexprUnfoldsTo [D1|P1]
\derives
   [D|P] \schemaEquals [D1|P1]
\end{zedrule}

\begin{zsection}
  \SECTION expand \parents unfold
\end{zsection}

\begin{zedrule}{SchemaToSet}
[D | P] = \{ D | P @ \theta [D | true] \}
\end{zedrule}

\end{document}
